### 2024-05-29 README.Rmd
I could get some more writing in today

The aim of the provisional guidelines was to inform the university community about the use of AI in "operational excellence". As we approach the Fall Term of 2024, with almost two years of cohabitation experience with AI for the masses at our backs, the time has come to update these guidelines. Pertinently, the institution, via its Office of the Provost, [launched a committee](#ai-committee-announced) to explore the use of AI across campus. This committee is composed of three sub-committees tasked with examining the use of AI in 1) teaching and learning, 2) research, and 3) operational excellence in "our work" (the Provost's words)(in all likelihood meaning the university's work; not sure if the Provost herself uses AI in addition to the majestic plural).

@@

Notice that this statement (like the definition of the work before the panels), takes for granted that 1) AIs are powerful tools, and 2) universities can be leading players in their adoption. This attitude is almost fatalistic^[In the sense of "fated to"] in its enthusiasm, and perfectly consistent with the hype observed around AI this time^[The pattern of cyclical hypes around AI even has a name: [AI winters](https://en.wikipedia.org/wiki/AI_winter); the current one with chatbots in the pole position is just the most recent of several hype cycles [see @siegel_ai_2023]] [@angwin_press_2024].

@@

The public was promised that GenAI would "accelerate sustainability" [@nakagawa_accelerating_2023]; that LLMs would transform sectors as diverse as health care, finance, human resources, insurance, "and beyond" by "automating customer self-service, accelerating response times on an increasing number of tasks as well as providing greater accuracy, enhanced routing and intelligent context gathering" [@ibm_what_2023]; according to Allerin^[Curiously, Allerin is not identified as an artificial intelligence company in @joshi_5_2022 but only in the much later article by @garsten_what_2024.], AI-powered self-driving cars would reduce accidents, car ownership, pollution and noise, and would automate logistics and make parking tickets obsolete [@joshi_5_2022; @garsten_what_2024]. In mental health care, chatbots could complement human therapists "via comprehensive data access and [analysis of] behavioral patterns" to "mimic practitioner questions and subsequently make recommendations based on a user’s inputs" [@silva_4_2023].

AI would even write code, perhaps helping itself to some self-programming (AI self-help?) as an outcome.

@@

```{r ai-writing-code, out.width="50%", fig.cap="\\label{fig:ai-bug-fix}Figure X. Example of an AI writing code"}
knitr::include_graphics("images/AI-bug-fix.png")
```

The list of promises has grown to encompass pretty much everything, from saving the planet all the way to micromanaging intimate facets of individual lives, for example by revolutionizing dating through the use of chatbots to help "people connect and find potential partners" (who presumably also use chatbots to do the same) [@khalatian_matchmaking_2023].

Higher-education was not spared the onslaught of promises.

@@

## What can GenAI tools do for students?{#what-can-ais-do-for-students}

@@

## What do GenAIs do for the mega-corporations pushing these tools? {#what-do-ais-to-corporations}

@@

## What do GenAIs to to everyone else? {#what-do-ais-do-to-everyone-else}

@@

Training data, much like technology, are not politically neutral, and several processes contribute to reproduce malignant biases, A way to explain this is by means of the concept of _platformed racism_ [@matamoros-fernandez_platformed_2017]. Platformed racism is the process whereby the functionalities of platforms are exploited to create and spread racist ideology (liking and sharing), often in a grass-roots fashion, but more ominously with the assist of bad-faith actors^[Online Russian disinformation has had a wide reach [@treyger_russian_2022] and furthermore is known to seed distrust between ethnic communities [@svedkauskas_analysis_2020]. As the U.S. Department of State Reports, Russian intelligence is known to exploit "all kinds of separatism and ethnic, social and racial conflicts, actively supporting all dissident movements--extremist, racist, and sectarian groups" to destabilize internal U.S. politics [@us_department_of_state_gec_2024, p. 51].]. Given how cagey corporations are about the data they use to train their models, it is anyone's guess how much dis- and misinformation has been passed to their AIs (this is part of the Unfathomable Data problem discussed by @bender_dangers_2021). In this way, AIs, rather than contributing to online safety, may end up amplifying the legacy of Russian interference [@svedkauskas_analysis_2020].]. The second aspect of platformed racism is the implicit endorsement of racist ideology through vague standards and [arbitrary moderation practices](https://www.bbc.com/news/technology-58058428) that exclude some views while allowing others to thrive [@myers_west_censored_2018; @hawkins_race_2023].

>>>>

The aim of the provisional guidelines was to inform the university community about the use of AI in "operational excellence". As we approach the Fall Term of 2024, with almost two years of cohabitation experience with AI for the masses at our backs, the time has come to update these guidelines. Pertinently, the institution, via its Office of the Provost, [launched a committee](#ai-committee-announced) to explore the use of AI across campus. This committee is composed of three sub-committees tasked with examining the use of AI in 1) teaching and learning, 2) research, and 3) operational excellence in "our work".

@@

Notice that this statement (like the definition of the work before the panels above), takes for granted that 1) AIs are powerful tools, and 2) universities can be leading players in their adoption. This attitude is almost fatalistic^[In the sense of "fated to"] in its enthusiasm, and perfectly consistent with the hype observed around AI this time^[The pattern of cyclical hypes around AI even has a name: [AI winters](https://en.wikipedia.org/wiki/AI_winter); the current one with chatbots in the pole position is just the most recent of several hype cycles [see @siegel_ai_2023]] [@angwin_press_2024].

@@

The public was promised that GenAI would "accelerate sustainability" [@nakagawa_accelerating_2023]; that LLMs would transform sectors as diverse as health care, finance, human resources, insurance, "and beyond" by "automating customer self-service, accelerating response times on an increasing number of tasks as well as providing greater accuracy, enhanced routing and intelligent context gathering" [@ibm_what_2023]; according to Allerin^[Curiously, Allerin is not identified as an artificial intelligence company in @joshi_5_2022 but only in the much later article by @garsten_what_2024.], AI-powered self-driving cars would reduce accidents, car ownership, pollution and noise, and would automate logistics and make parking tickets obsolete [@joshi_5_2022; @garsten_what_2024]. But there is no reason to stop at self-driving vehicles: McMaster University's Canada Excellence Research Chair Laureate Ali Emadi "hopes to help revolutionize and influence everything from personal travel to city planning" by putting AI directly into our transportation systems [@dorey_driving_2023]. In mental health care, chatbots could complement human therapists "via comprehensive data access and [analysis of] behavioral patterns" to "mimic practitioner questions and subsequently make recommendations based on a user’s inputs" [@silva_4_2023].

AI would even write code, perhaps helping itself to some self-programming along the way: AIs that write themselves, or AI self-help?

@@

```{r ai-writing-code, out.width="50%", fig.cap="\\label{fig:ai-bug-fix}Figure X. An example of Atlassian AI writing code"}
knitr::include_graphics("images/AI-bug-fix.png")
```

The list of promises about what AI will do has grown to encompass pretty much everything, from saving the planet from ourselves, all the way down to micromanaging intimate facets of individual lives, for example by revolutionizing dating through the use of chatbots to help "people connect and find potential partners" (who presumably also use chatbots to do the same) [@khalatian_matchmaking_2023].

Higher-education has not spared the onslaught of promises.

@@

## What do GenAI tools do for students?{#what-can-ais-do-for-students}

@@

## What do GenAIs do for the mega-corporations pushing them? {#what-do-ais-to-corporations}

@@

To read:

@sriram_ai_2023

## What do GenAIs do to everyone else? {#what-do-ais-do-to-everyone-else}

@@

Training data, much like technology, are not politically neutral, and several processes contribute to reproduce malignant biases, A way to explain this is by means of the concept of _platformed racism_ [@matamoros-fernandez_platformed_2017]. Platformed racism is the process whereby the functionalities of platforms are exploited to create and spread racist ideology (liking and sharing), often in a grass-roots fashion, but more ominously with the assist of bad-faith actors^[Online Russian disinformation has had a wide reach [@treyger_russian_2022] and furthermore is known to seed distrust between ethnic communities [@svedkauskas_analysis_2020]. As the U.S. Department of State Reports, Russian intelligence is known to exploit "all kinds of separatism and ethnic, social and racial conflicts, actively supporting all dissident movements--extremist, racist, and sectarian groups" to destabilize internal U.S. politics [@us_department_of_state_gec_2024, p. 51].]. Given how cagey corporations are about the data they use to train their models, it is anyone's guess how much dis- and misinformation has been passed to their AIs (this is part of the Unfathomable Data problem discussed by @bender_dangers_2021). In this way, AIs, rather than contributing to online safety, may end up amplifying the legacy of Russian interference [@svedkauskas_analysis_2020].]. The second aspect of platformed racism is the implicit endorsement of racist ideology through vague standards and arbitrary moderation practices^[According to the Center for Countering Digital Hate, platforms like facebook, Instagram, Tiktok, Twitter, and YouTube, failed to act 84% of the time on hundreds of posts that had millions of views, with facebook being the worst performer [@center_for_countering_digital_hate_failure_2021].] that exclude some views while allowing others to thrive [@myers_west_censored_2018; @hawkins_race_2023].
